---
title: "AVISO+ Velocity Dataset"
author: "Sebastian Di Geronimo"
date: 2025-07-14
format: html
editor: source
---

# ---- Summary of Document ----

Download AVISO+ data to get geostrophic velocities in the Florida Keys. The 
geostrophic velocities can be used to estimate where the Florida Current (FC) is
in relation to the coastline. The years used are 2015 - Nov 2024 (based on 
data availability).

A threshold of >= 0.5 m/s is used to determine the FC based on a Zhang et al., 2022.


## ---- AVISO+ Sea Surface Height and Velocity ----

Data downloaded from `Copernicus` for both absolute dynamic topography (m) and
north/east surface velocities (m/s)

Data:       Global Ocean Gridded L4 Sea Surface Heights And Derived Variables
Search:     <https://data.marine.copernicus.eu/products>
Product ID: `SEALEVEL_GLO_PHY_L4_MY_008_047`
Dataset ID: `cmems_obs-sl_glo_phy-ssh_my_allsat-l4-duacs-0.125deg_P1D`
Description: <https://data.marine.copernicus.eu/product/SEALEVEL_GLO_PHY_L4_MY_008_047/description>
Bounds:     
  Lon: -83.8125, -78.3125  
  Lat:  22.8125,  26.4375
Note: Select all data products when downloading from `MyOceanPro`
      If downloading from `description` page, go to `Data access` then `subset`, 
      then `automate`. Copy the `automate` and modify script below to download
      multiple dates.

Data Info:
<https://documentation.marine.copernicus.eu/PUM/CMEMS-SL-PUM-008-032-068.pdf>


Velocites extraction inspired from:
<https://help.marine.copernicus.eu/en/articles/9711615-how-to-plot-current-vectors-using-r>


## Download Data from `Copernicusmarine` App

- using `copernicusmarine.exe` File

From: <https://help.marine.copernicus.eu/en/articles/10750437-copernicus-marine-toolbox-executable-no-installation>

I recommend moving the file to this directory where this project is stored

To start, open `cmd` or `PowerShell` from explorer. `cd` to location where `.exe`
is located. Then run `copernicusmarine login` and enter "Username" and
"Password". Once completed, you should be able to run the download




variables:

adt            - absolute dynamic topography [m]
tpa_correction - instrument drift correction [m]
flag_ice       - ice flag
sla            - sea level anomaly [m]
err_sla        - sea level anomaly error [m]

u - eastward
  ugos           - vector velocity [m/s]
  ugosa          - vector velocity anomly [m/s]
  err_ugosa      - vector velocity error [m/s]
  
v - northward
  vgos           - vector velocity [m/s]
  vgosa          - vector velocity anomly [m/s]
  err_vgosa      - vector velocity error [m/s]


# ---- Setup ----


## Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )

# source aviso functions
source(here("R", "aviso_functions.R"))

# set paths to downloaded data, ffmpeg, and animation
loop <- FALSE
loop <- TRUE
if (loop) {
  message("Adding loop current")
  add_loop <- "loop_current"
  xmin <- -89
} else {
  add_loop <- ""
  xmin <- -83.8125
}


path_save <- here("data", "raw", "copernicus", add_loop) 

dir_create(path_save)

# path to executable and path to save data
path_copernicusmarine <- here("copernicusmarine.exe")
```


### Setup Constants used in Download

```{r download-aviso-setup}
# set date range by month
months <-
  seq(as_date("2015-01-01"),
      as_date("2024-12-31"),
      by = "1 month")
# months <-
#   seq(as_date("2015-01-01"),
#       as_date("2015-01-31"),
#       by = "1 month")

# set spatial extent
aviso_bbox <-
  c(
    xmin = xmin,
    xmax = -78.3125,
    ymin = 22.8125,
    ymax = 26.4375
  )

# select vars
vars <-
  c(
    "adt",
    "err_sla",
    "err_ugosa",
    "err_vgosa",
    "flag_ice",
    "sla",
    "tpa_correction",
    "ugos",
    "ugosa",
    "vgos",
    "vgosa"
  )

cat("Date Range: ", format(months[c(1, length(months))], "%B %Y"), sep = "\n- ")

```

### Download AVISO+ Data

```{r download-aviso}
# set to TRUE if downloading data
download <- FALSE
# download <- TRUE

# set to TRUE if want to overwrite previous data
overwrite <- FALSE
# overwrite <- TRUE

aviso_download(
  months, 
  vars, 
  bounds = aviso_bbox,
  path_copernicusmarine,
  path_save = path_save,
  download,
  overwrite
)

# path_box <- rstudioapi::selectDirectory()
path_box <- here("C:/Users/sebastian15/Box/aviso_velocity/data/raw/copernicus", 
                 add_loop)

dir_ls(path_save1) %>% 
file_copy(path_box)
```

