---
title: "HYCOM-TSIS Hourly Download"
author: "Sebastian Di Geronimo"
date: 2026-08-19
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

## Purpose

This document is used to download HYCOM-TSIS 1/100º Gulf of Mexico Reanalysis 
at a high temporal resolution (i.e. hourly) data for sepcified date ranges.

This will be used in particle tracking at hourly resolution instead of once a 
day which was previously done.

## HYCOM

Host Website: <https://www.hycom.org/>
Catalog: <https://ncss.hycom.org/thredds/catalog.html>
Forum for HYCOM: <https://groups.google.com/a/hycom.org/g/forum>
- help documents

## HYCOM-TSIS 1/100º Gulf of Mexico Reanalysis

<https://www.hycom.org/data/gomb0pt01/gom-reanalysis>

Uses Tendral Statistical Interpolation (T-SIS) package to get high resolution
(Srinivasan et al., 2022, www.tendral.com/tsis). multivariate linear statistical 
estimation given a predicted ocean state and observations

Assimilates
- SLA (T/P, JASON 1 and 2, Envisat, GFO and Cyrosat)
- SST (GHRSST and AVHRR)
- Temp/salinity from ARGO floats


Hourly 2001-01-01 to "Present" (2-3 month delay)
- 40 depth layers
- velocity in u, v 
- vertical velocity in w

For addtional documentation:
<https://www.hycom.org/hycom/documentation>


## HYCOM and NCODA Gulf of Mexico 1/25° Analysis
This is some information about a higher resolution model from HYCOM and NCODA
<https://portal.secoora.org/#module-metadata/8bc4c4e7-ddbb-462e-a07c-c0e0d5fbeb3f>

Hourly from 2017 - Jan 2023 at 1/25°
Variables:
- sea water velocity
- water surface elevation
- salinity
- water temperature


# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  ncdf4, sf, metR, tidyterra, oce
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(),
  tidyr::extract()
  )
  
shelf(terra)

source(here("R", "aviso_functions.R"))
```

## 2.2 Cruise Dates

```{r ctd-dates}
# hour to extract 0 - 23
hours_of_day <- c(12)
hours_of_day <- c(0:23)
# hours_of_day <- c(0, 6, 12, 18)

traj_req_loc <-
  here(cloud_dir, "hycom_vid_img", "particle_tracker") %>% 
  dir_ls(regexp = "trajectory_request") %>% 
  str_subset("~", negate = TRUE) %>%
  readxl::read_xlsx(sheet = "Backward_Keys_11") %>% 
  select(cruise_id, sampling_start2, sampling_end2, particle_initial_time_utc) %>%
  mutate(
    .by = cruise_id,
    dwnld_strt = min(sampling_start2),
    dwnld_end = min(sampling_end2),
  ) %>%
  filter(
    # cruise_id == "WS1018"
  ) %>%
  print()

end_date <- any(traj_req_loc$dwnld_strt >= ymd_hms("2024-09-01 0:0:0"))


# years to search
srch_year <- 
  unique(c(year(traj_req_loc$dwnld_strt), year(traj_req_loc$dwnld_end)))

all_dates <-
  map2(
    traj_req_loc$dwnld_strt,
    traj_req_loc$dwnld_end,
    \(.x, .y) {
      seq.POSIXt(.x, .y, "days")
    }
  ) %>%
  list_c() %T>% 
  print()

```



# 3.0 ---- Download HYCOM-TSIS Data ----

Info: <https://www.hycom.org/data/gomb0pt01/gom-reanalysis>
From: <https://tds.hycom.org/thredds/catalog/datasets/GOMb0.01/reanalysis/data/catalog.html>

Help for download flags: <https://docs.unidata.ucar.edu/tds/current/userguide/ncss_grid.html>
- dateTime has the form: '-'? yyyy '-' mm '-' dd 'T' hh ':' mm ':' ss ('.' s+)? (zzzzzz)?

Base URL:
https://ncss.hycom.org/thredds/ncss/datasets/GOMb0.01/reanalysis/data/YYYY/xxx_archv.YYYY_DDD_HH_NN.nc

Experiment Numbers:
This part `xxx_archv.YYYY_DDD_HH_NN.nc` is described here
xxx: experiment number (i.e. 010, 023 or 026 for date ranges)
YYYY: year
DDD: day (julian day)
HH: hour
NN: Netcdf type (i.e. 2d (surface) or 3z (with depth))

HYCOM-TSIS GOMb0.01:
010_archv.YYYY_DDD_HH_NN.nc: 2001_001_00 to 2017_152_18
023_archv.YYYY_DDD_HH_NN.nc: 2017_152_19 to 2024_032_18
026_archv.YYYY_DDD_HH_NN.nc: 2024_001_19 to *PRESENT*


Parameters used in link for 3z:
- variables
  - var=salinity&
  - var=u&
  - var=v&
  - var=w_velocity&
  - var=water_temp
- lat/lon
  - north=26.4375&
  - west=-83.8125&
  - east=-78.3125&
  - south=22.8125
- horizStride=1
  - number of data points to skip along horiztonal (x and y)
- vertStride=1
  - number of data points to skip along vertical (z)
- time=2017-01-02T10%3A01%3A52.500Z
  - yyyy-mm-ddThh:mm::ss
  - %3A is : (colon in URL encoding)
- vertCoord=0.0
  - single level 0.0, 2.0, 4.0, ..., 5000 m

- accept=netcdf4
  - type of netcdf (netcdf or netcdf4)


Depth options (40): 
  0.0 2.0 4.0 6.0 8.0 10.0 12.0 15.0 20.0 25.0 30.0 35.0 40.0 45.0 
  50.0 60.0 70.0 80.0 90.0 100.0 125.0 150.0 200.0 250.0 300.0 350.0 400.0 
  500.0 600.0 700.0 800.0 900.0 
  1000.0 1250.0 1500.0 2000.0 2500.0 3000.0 4000.0 5000.0


## 3.1 Constants

Need to select the correct url for the time period being selected.

`GOMb0.01/reanalysis/data/` - 2001-01-01 - 2024-09-01 18:00

`GOMe0.01/expt_02.8/data/archive/` - 2024-09-02 19:00 - present

```{r constants}
# ---- catalog url ---- #
# 2001-01-01 - 2024-08-31
hycom_url <- "https://tds.hycom.org/thredds/catalog/datasets/GOMb0.01/reanalysis/data/"
suffix <- "GOMb0.01"

if (end_date) {
  # when any dates are above what is possible with original set of hycom data
  # 2024-09-01 -
  hycom_url <- "https://tds.hycom.org/thredds/catalog/datasets/GOMe0.01/expt_02.8/data/archive/"
  all_dates <- all_dates[all_dates >= ymd_hms("2024-09-01 0:0:0")]
  suffix <- "GOMe0.01"
}

# ---- constants ---- #
vert    <- format(0, nsmall = 1) # depth
horizon <- 1 # number of data points to skip, this save data space

lat_lon <- list(
  # long
  "xmin" = -83.8125,
  "xmax" = -78.3125,
  #lat
  "ymin" = 22.8125,
  "ymax" = 26.4375
)
```

## 3.2 Search HYCOM For Data Files

This search only needs to happen once to get the file name. 

To subset the data, change the info in constants 

```{r search-hycom}
if (exists("cloud_dir")) {
  hycom_info <-
    here(
      cloud_dir,
      "data",
      # "hycom_100deg_download_instructions_3.csv"
      glue("hycom_100deg_download_{nrow(traj_req_loc)}_cruises_{length(hours_of_day)}_times_{suffix}.csv")
      )
} else {
  hycom_info <- 
    here(
    rstudioapi::selectDirectory(),
    # "hycom_100deg_download_instructions_3.csv"
    glue("hycom_100deg_download_{nrow(traj_req_loc)}_cruises_{length(hours_of_day)}_times_{suffix}.csv")
  )
}

# TODO: replace search with function `hycom_search()`
if (!file_exists(hycom_info)) {
  shelf(rvest)
  # specific url base for searching the database
  specifics <-
    paste0(
      "var=salinity&var=u&var=v&var=w_velocity&var=water_temp&",
      "north={lat_lon$ymax}&west={lat_lon$xmin}&east={lat_lon$xmax}&south={lat_lon$ymin}&",
      "disableProjSubset=on&",
      "horizStride={horizon}&",
      "time={date}&",
      "vertCoord={vert}&accept=netcdf4"
    )

  # ---- web searching ---- #

  message("This will take a while!")
  
  # ---- search URL for every file ---- #
  # selection based on minimum year and hour od day
  message("1. Searching for every file name")
  
  search_individual_files <-
    paste0(hycom_url, "catalog.html") %>%
    read_html() %>%
    html_elements("a tt") %>%
    html_text() %>%
    str_subset("\\d{4}") %>%
    str_remove("/") %>%
    as.numeric() %>%
    tibble(file = hycom_url, year = .) %>%
    
    # filter year
    filter(year %in% srch_year) %>% 
    
    mutate(file_yr = glue("{file}{year}/")) %>%
    mutate(
      file_day = map(
        file_yr,
        \(.file) {
          x <- 
            paste0(.file, "catalog.html") %>%
            read_html() %>%
            html_element("body") %>%
            html_table() %>%
            pull(1) %>%
            str_subset("_3z\\.nc") %>%
            tibble(file = .file, day = .) %>%
            mutate(
              day_yr = str_extract(day, "\\d{4}_\\d{3}"),
              day_yr = as_date(day_yr, format = "%Y_%j"),
              hour = str_extract(day, "(\\d{2})_3z\\.nc", group = 1),
              hour = as.numeric(hour)
            ) %>%
            filter(
              day_yr %in% as_date(all_dates),
              hour %in% hours_of_day
            ) %>%
            mutate(
              part_url = str_extract(file, "datasets/GO.*"),
              file_day = paste0(file, "catalog.html?dataset=", part_url, day)
            )
        }
      )
    ) %T>%
    print()


  
  # ---- search the file URL for download ---- #
  message("2. Searching for file location")
  extract_file_func <-  function(.path) {
    print(basename(.path))
    .path %>%
      read_html() %>%
      html_element("body") %>%
      html_element("ol") %>%
      html_text() %>%
      str_extract("NetcdfSubset.*\\.nc") %>%
      str_remove("NetcdfSubset:.*//") %>%
      paste0("https://", ., "/dataset.html")
  }

  extract_file_location <-
    search_individual_files %>%
    unnest(file_day, names_sep = "_") %T>%
    print() %>%
    mutate(
      info = map_chr(
        file_day_file_day,
        possibly(
          extract_file_func, 
          otherwise = NA,
          quiet = FALSE
          )
        )
      ) %T>%
    print()
   
  # ---- extract date from file URL ---- #
  message("3. Extracting date from URL")
  message(glue("A total of {nrow(extract_file_location)} day(s) to be searched."))
  extract_date <- function(.path) {
    print(basename(dirname(.path)))
    .path %>%
      read_html() %>%
      html_element("body") %>%
      html_elements("h3 span") %>%
      html_text() %>%
      str_subset("\\d{4}.*") %>%
      str_replace_all(":", "%3A")
    }
  
  
  extract_date_location <-
    extract_file_location %>%
    mutate(
      date = map(
        info,
        possibly(
          extract_date,
          otherwise = NA,
          quiet = FALSE
        )
      )
    ) %>%
    unnest(date) %T>%
    print()

  # ---- setup final url before adding constants ---- #
  message("4. Final setup of date, file path and file name")
  final_download_info <-
    extract_date_location %>%
    mutate(
      .keep = "used",
      date,
      final_path =
        str_remove(info, "grid/") %>%
          str_replace("/dataset.html", "?"),
      final_path = paste0(
        final_path,
        specifics
      ),
      hour = str_extract(info, "_(\\d{2})_3z\\.nc", group = 1),
      file_name = paste0("hycom_", str_remove(date, "T.*"), "_H", hour, ".nc4"),
      file_name = str_replace_all(file_name, "-", "_")
    )  %>%
    select(-hour) %T>%
    print()
  
  # ---- write final info to file to save time ---- #
  message(glue("5. Write instructions to {hycom_info}"))
  write_csv(final_download_info, hycom_info)
  
  unshelf(rvest)
  
} else {
  message(
    glue("HYCOM search previously completed on ", 
         "{file_info(hycom_info)$birth_time}")
    )
   
  final_download_info <- read_csv(hycom_info, show_col_types = FALSE)
}

```




## 3.3 Download HYCOM-TSIS Data

Find file with all URLs to HYCOM Data

```{r download-hycom}
if (exists("cloud_dir")) {
  dwnld_path <- here(cloud_dir, "data", "raw", "hycom_cruise_select")
} else {
  dwnld_path <- here("data", "raw", "hycom_cruise_select")
}

# apply final search parameters
path_save <- 
  final_download_info %>%
  mutate(
    year = ymd(str_extract(date, "\\d{4}-\\d{2}-\\d{2}")),
    year = year(year),
    file_name = paste0(year, "/", file_name)
  ) %>% 
  mutate(
    final_path = map2_chr(
      final_path,
      date,
      \(final, date) {
        glue(final)
      }
    )
  ) %T>%
  print() 

# create folder for each year month pair
path_save %>% 
 distinct(year) %$%
  walk(
    year,
    \(.x) {
      here(dwnld_path, .x) %>% 
        dir_create()
    }
  )
```

### 3.3.2 Download HYCOM Data One at a Time

```{r single-download}
path_save %>% 
  mutate(
    date2 = str_remove(date, "T.*") %>% ymd(),
    rows = row_number(),
    max_r = nrow(path_save)) %>% 
  
  # filter dates
  filter() %T>%
  {
    
    message(paste("Downloading", nrow(.), "files"))
    message(paste("Date ranges:", paste(range(.$date2), collapse = " - ")))
  } %$%
  pwalk(
    list(
      final_path,
      file_name,
      rows, 
      max_r
      ),
    \(.url, .file_name, idx, max_row) {
      message(paste(idx, "of", max_row))
      message(paste("HYCOM URL:", .url))
      message(paste("\nDate:", ymd(str_extract(basename(.file_name), "\\d+_\\d+_\\d+")), "\n"))
      message(paste("Save Path:", dirname(here(dwnld_path, .file_name))))
      message(paste("File name:", basename(.file_name), "\n"))
      
      if (file_exists(here(dwnld_path, .file_name))) {
         message("File exists. Skipping!\n")
         message("---\n")
         return()
      }
      # message("---\n")
      message("Downloading\n")
      download.file(
        url      = .url,
        destfile = here(dwnld_path, .file_name),
        method   = "curl"
      )
      message("\nSuccesful download!")
      message("\n---\n")
    }
  )

# check file size is less than than 1 Kb, then delete 
here(dwnld_path) %>%
  dir_ls(regexp = "\\.nc4", recurse = TRUE) %>%
  file_info() %>%
  dplyr::arrange(size) %>%
  dplyr::filter(size < fs_bytes("1Kib")) %>%
  pull(path)  %T>% 
  print() %>%
  {
    if (length(.) > 0 && menu(c("Keep small files", "Delete")) == 2) {
      file_delete(.)
    }
  }
  
```

```{r}
here(dwnld_path) %>%
  dir_ls(regexp = "\\.nc4", recurse = TRUE) %>% 
  tibble(file = .) %>%
  # slice(1:5) %>%
  mutate(
    year = str_extract(file, "\\d{4}"),
    yrmth = str_extract(file, "\\d{4}_\\d{2}"),
    yrmth = str_replace(yrmth, "_", "/")
  ) %>%
  count(year) %T>% 
  print() 
```
