---
title: "Modeled Currents"
author: "Sebastian Di Geronimo"
date: 2025-08-06
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

Host Website: <https://www.hycom.org/>
Catalog: <https://ncss.hycom.org/thredds/catalog.html>
Forum for HYCOM: <https://groups.google.com/a/hycom.org/g/forum>
- help documents

## HYCOM-TSIS 1/100º Gulf of Mexico Reanalysis

<https://www.hycom.org/data/gomb0pt01/gom-reanalysis>

Uses Tendral Statistical Interpolation (T-SIS) package to get high resolution
(Srinivasan et al., 2022, www.tendral.com/tsis). multivariate linear statistical 
estimation given a predicted ocean state and observations

Assimilates
- SLA (T/P, JASON 1 and 2, Envisat, GFO and Cyrosat)
- SST (GHRSST and AVHRR)
- Temp/salinity from ARGO floats


Hourly 2001-01-01 to "Present" (2-3 month delay)
- 40 depth layers
- velocity in u, v 
- vertical velocity in w

For addtional documentation:
<https://www.hycom.org/hycom/documentation>


## HYCOM and NCODA Gulf of Mexico 1/25° Analysis
This is some information about a higher resolution model from HYCOM and NCODA
<https://portal.secoora.org/#module-metadata/8bc4c4e7-ddbb-462e-a07c-c0e0d5fbeb3f>

Hourly from 2017 - Jan 2023 at 1/25°
Variables:
- sea water velocity
- water surface elevation
- salinity
- water temperature


## CNAPS

Coupled Northwest Atlantic Prediction System
May 1, 2016 20:00 (EDT) - Jul 2, 2019 17:00 (EDT)
3 hours
<https://portal.secoora.org/#metadata/4bc223e1-6b14-46f5-9832-f8e9ff061a36/13cee4ae-2d36-454c-a6bc-1dcc223d819c>
Variables:
- northward velocity
- bottom U-momentum stress
- bottom V-momentum stress
- currents
- eastward velocity
- evaporation rate
- free surface
- net latent heat flux
- net longwave radiation flux
- net sensible heat flux
- potential temperature
- rain fall rate
- salinity
- solar shortwave radiation flux
- surface net heat flux
- surface net salt flux
- surface U-momentum stress
- surface V-momentum stress
- vertical momentum component
- wind induced mean wavelength
- wind induced significant wave height
- wind induxed wave direction


## SABGOM
South Atlantic Bight and Gulf of Mexico
Apr 29, 2016 20:00 (EDT) - Dec 25, 2017 19:00 (EST)
3 hours
<https://portal.secoora.org/#metadata/9cf0615d-948a-46f8-ad4b-73c887b9eaa3/ab6cb99f-ed82-46cd-a0c7-cf1b2069adfc>
incorporating the ROMS model for circulation, the WRF model for atmospheric 
circulation, the SWAN model for surface waves




# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  ncdf4
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(),
  tidyr::extract()
  )
  

```

## 2.2 Cruise Dates

```{r ctd-dates}
source(here("R", "aviso_functions.R"))
# minimum year to extract from
min_year <- 2015
hours_of_day <- 12

# previous days to load from last CTD data
prev_day <- 14

if (exists("cloud_dir")) {
  min_max_date <-
    here(cloud_dir, "data", "raw") %>%
    dir_ls(regexp = "ctd_min_max_date")  %>%
    read_csv(show_col_types = FALSE) %>% 
    mutate(
      year = year(max_time)
    )  %T>%
    print()
} else {
  min_max_date <-
    rstudioapi::selectFile() %>%
    read_csv(show_col_types = FALSE) %>% 
    mutate(
      year = year(max_time)
    ) %T>%
    print()
} 


cruise_ids <- c("WS16004", "WS17030", "WS17212", "WS18008", "WS20342", "WS21212")

min_max_date <- 
  min_max_date %>% 
  filter(year >= min_year) %>% # start of HYCOM data
  # filter(year %in% c(2016:2018)) %>%
  # filter(cruise_id %in% cruise_ids) %>% 
  # slice(1:3) %>% 
  pull(max_time) %>% 
  
  # add all dates from max date minus x days
  map(
    \(.x) {
      as_date((.x-days(prev_day)):.x)
    }
  ) %>% 
  list_c() %T>% 
  print()
```


# 3.0 Download HYCOM-TSIS Data 

Info: <https://www.hycom.org/data/gomb0pt01/gom-reanalysis>
From: <https://tds.hycom.org/thredds/catalog/datasets/GOMb0.01/reanalysis/data/catalog.html>

Help for download flags: <https://docs.unidata.ucar.edu/tds/current/userguide/ncss_grid.html>
- dateTime has the form: '-'? yyyy '-' mm '-' dd 'T' hh ':' mm ':' ss ('.' s+)? (zzzzzz)?

Base URL:
https://ncss.hycom.org/thredds/ncss/datasets/GOMb0.01/reanalysis/data/YYYY/xxx_archv.YYYY_DDD_HH_NN.nc

Experiment Numbers:
This part `xxx_archv.YYYY_DDD_HH_NN.nc` is described here
xxx: experiment number (i.e. 010, 023 or 026 for date ranges)
YYYY: year
DDD: day (julian day)
HH: hour
NN: Netcdf type (i.e. 2d (surface) or 3z (with depth))

HYCOM-TSIS GOMb0.01:
010_archv.YYYY_DDD_HH_NN.nc: 2001_001_00 to 2017_152_18
023_archv.YYYY_DDD_HH_NN.nc: 2017_152_19 to 2024_032_18
026_archv.YYYY_DDD_HH_NN.nc: 2024_001_19 to *PRESENT*


Parameters used in link for 3z:
- variables
  - var=salinity&
  - var=u&
  - var=v&
  - var=w_velocity&
  - var=water_temp
- lat/lon
  - north=26.4375&
  - west=-83.8125&
  - east=-78.3125&
  - south=22.8125
- horizStride=1
  - number of data points to skip along horiztonal (x and y)
- vertStride=1
  - number of data points to skip along vertical (z)
- time=2017-01-02T10%3A01%3A52.500Z
  - yyyy-mm-ddThh:mm::ss
  - %3A is : (colon in URL encoding)
- vertCoord=0.0
  - single level 0.0, 2.0, 4.0, ..., 5000 m

- accept=netcdf4
  - type of netcdf (netcdf or netcdf4)


Depth options (40): 
  0.0 2.0 4.0 6.0 8.0 10.0 12.0 15.0 20.0 25.0 30.0 35.0 40.0 45.0 
  50.0 60.0 70.0 80.0 90.0 100.0 125.0 150.0 200.0 250.0 300.0 350.0 400.0 
  500.0 600.0 700.0 800.0 900.0 
  1000.0 1250.0 1500.0 2000.0 2500.0 3000.0 4000.0 5000.0


## 3.1 Constants
```{r constants}
# ---- catalog url ---- #
hycom_url <- "https://tds.hycom.org/thredds/catalog/datasets/GOMb0.01/reanalysis/data/"

# ---- constants ---- #
vert    <- format(0, nsmall = 1) # depth
horizon <- 1 # number of data points to skip, this save data space

# lon
xmin    <- -83.8125 
xmax    <- -78.3125
# lat
ymin    <- 22.8125
ymax    <- 26.4375
```

## 3.2 Search HYCOM For Data Files

This search only needs to happen once to get the file name. 

To subset the data, change the info in constants 

```{r search-hycom}
if (exists("cloud_dir")) {
  hycom_info <-
    here(
      cloud_dir,
      "data",
      "hycom_100deg_download_instructions.csv"
      )
} else {
  hycom_info <- 
    here(
    rstudioapi::selectDirectory(),
    "data",
    "hycom_100deg_download_instructions.csv"
  )
}


if (!file_exists(hycom_info)) {
  shelf(rvest)
  # specific url base for searching the database
  specifics <-
    paste0(
      "var=salinity&var=u&var=v&var=w_velocity&var=water_temp&",
      "north={ymax}&west={xmin}&east={xmax}&south={ymin}&",
      "disableProjSubset=on&",
      "horizStride={horizon}&",
      "time={date}&",
      "vertCoord={vert}&accept=netcdf4"
    )


  # ---- web searching ---- #

  # ---- search URL for every file
  # selection based on minimum year and hour od day
  search_individual_files <-
    paste0(hycom_url, "catalog.html") %>%
    read_html() %>%
    html_elements("a tt") %>%
    rvest::html_text() %>%
    str_subset("\\d{4}") %>%
    str_remove("/") %>%
    as.numeric() %>%
    tibble(file = hycom_url, year = .) %>%
    filter(year >= min_year) %>% # filter year
    mutate(file_yr = glue("{file}{year}/")) %>%
    mutate(
      file_day = map(
        file_yr,
        \(.file) {
          paste0(.file, "catalog.html") %>%
            read_html() %>%
            html_element("body") %>%
            html_table() %>%
            pull(1) %>%
            str_subset(glue("{hours_of_day}_3z\\.nc")) %>%
            tibble(file = .file, day = .) %>%
            mutate(
              part_url = str_extract(file, "datasets/GO.*"),
              file_day = paste0(file, "catalog.html?dataset=", part_url, day)
            )
        }
      )
    ) %T>%
    print()

  # ---- search the file URL for download
  extract_file_location <-
    search_individual_files %>%
    unnest(file_day, names_sep = "_") %T>%
    print() %>%
    # slice(1, .by = year) %>%
    mutate(
      info = map(
        file_day_file_day,
        \(.path) {
          .path %>%
            read_html() %>%
            html_element("body") %>%
            html_element("ol") %>%
            html_text() %>%
            str_extract("NetcdfSubset.*\\.nc") %>%
            str_remove("NetcdfSubset:.*//") %>%
            paste0("https://", ., "/dataset.html")
        }
      )
    ) %>%
    unnest(info) %T>%
    print()

  # ---- extract date from file URL
  extract_date_location <-
    extract_file_location %>%
    mutate(
      date = map(
        info,
        \(.path) {
          .path %>%
            read_html() %>%
            html_element("body") %>%
            html_elements("h3 span") %>%
            html_text() %>%
            str_subset("\\d{4}.*") %>%
            str_replace_all(":", "%3A")
        }
      )
    ) %>%
    unnest(date) %T>%
    print()

  # ---- setup final url before adding constants
  final_download_info <-
    extract_date_location %>%
    mutate(
      .keep = "used",
      date,
      final_path =
        str_remove(info, "grid/") %>%
          str_replace("/dataset.html", "?"),
      final_path = paste0(
        final_path,
        specifics
      ),
      file_name = paste0("hycom_", str_remove(date, "T.*"), ".nc4"),
      file_name = str_replace_all(file_name, "-", "_")
    ) %T>%
    print()
  
  # ---- write final info to file to save time
  write_csv(final_download_info, hycom_info)
  
  unshelf(rvest)
  
} else {
  
  final_download_info <- read_csv(hycom_info, show_col_types = FALSE)
}


```


## 3.3 Download HYCOM-TSIS Data




Find file with all URLs to HYCOM Data

```{r download-hycom}
if (exists("cloud_dir")) {
  dwnld_path <- here(cloud_dir, "data", "raw", "hycom")
} else {
  dwnld_path <- here("data", "raw", "hycom")
}

dir_create(dwnld_path)

# apply final search parameters
path_save <- 
  final_download_info %>%
  mutate(
    final_path = map2_chr(
      final_path,
      date,
      \(final, date) {
        glue(final)
      }
    )
  ) %T>%
  print()
```

### 3.3.2

```{r single-download}
path_save %>% 
  mutate(date2 = str_remove(date, "T.*") %>% ymd()) %$%
  # filter dates
  # filter(date2 %in% min_max_date) %$%
  # slice(1500:1502) %$%
  walk2(
    final_path,
    file_name,
    \(.url, .file_name) {
      print(.url)
      cat("\n\n\n")
      print(.file_name)
      
      if (file_exists(here(dwnld_path, .file_name))) {
         message("File exists. Returning.")
         return()
      }
      
      download.file(
        url      = .url,
        destfile = here(dwnld_path, .file_name),
        method   = "curl"
      )
    }
  )

```


### 3.3.3 Parallel Download

May parallel download data if have access to the `furrr` and `progressr`

```{r parallel-download}
if (FALSE) {
  shelf(progressr, furrr)
  plan(multisession, workers = 5)
  
  
  filt_download <-  
    # filter dates
    path_save %>%
    mutate(date2 = str_remove(date, "T.*") %>% ymd()) %T>%
    # filter(date2 %in% min_max_date) %$%
    # slice(1:500) %T>% 
    print()
  
  with_progress({
    
    p <- progressor(steps = nrow(filt_download))
    
    filt_download %$%
      future_walk2(
        final_path,
        file_name,
        \(.url, .file_name) {
          p()
          print(.url)
          cat("\n\n\n")
          print(.file_name)
           if (file_exists(here(dwnld_path, .file_name))) {
             message("File exists. Returning.")
             return()
           }
           
           download.file(
             url      = .url,
             destfile = here(dwnld_path, .file_name),
             method   = "curl"
           )
          },
          .progress = TRUE)
  })
}
```

# 4.0 Load Dataset

```{r load-data}
# local storage
hycom_path <- here("data", "raw", "hycom") %>%
  dir_ls() %T>% 
  print()

# cloud storage
hycom_path <- here(cloud_dir,"data", "raw", "hycom") %>%
  dir_ls() %T>% 
  {
  sample(., 10) %>% print()
  }

# hycom_path <- rstudioapi::selectFile(path = here("..", "misc_ideas", "data", "raw", "hycom"))

```


```{r}

shelf(terra)
```


```{r}
hycom_dat <- ncdf4::nc_open(hycom_path[10])

hycom_dat
names(hycom_dat)
hycom_dat$var %>%
  names()
names(hycom_dat$dim)

if (length(ncvar_get(hycom_dat, "Depth")) > 1) {
  # if multiple depths
  message("HYCOM data has multiple depths (m).")
  hycom_u <- ncvar_get(hycom_dat, "u")[, , 1]
  hycom_v <- ncvar_get(hycom_dat, "v")[, , 1]
} else {
  # if single depth
  message("HYCOM data has one depth (m).")
  hycom_u <- ncvar_get(hycom_dat, "u") 
  hycom_v <- ncvar_get(hycom_dat, "v") 
}

lat  <- ncvar_get(hycom_dat, "Latitude")
lon  <- ncvar_get(hycom_dat, "Longitude")

 # create grid data
      hycom_grid <-
        expand.grid(lon = lon, lat = lat) %>%
        mutate(
          # adt     = as.vector(adt), # absolute dynamic topography
          hycom_v    = as.vector(hycom_v), # north geostrophic velocity
          hycom_u    = as.vector(hycom_u), # east geostrophic velocity
          vel_mag = sqrt(hycom_v^2 + hycom_u^2) # calculate current velocity
        )

      # dat_grid_temp[[i]] <- ssh_grid

ncvar_get(hycom_dat, "MT") %>% 
  lubridate::as_date(origin = "1900-12-31")
  

nc_close(hycom_dat)
```

```{r}
list(hycom_grid) %>%
  rast(crs = "EPSG:4326")

hycom_grid %>%
      
  rast()

unique(hycom_grid$lon)
unique(hycom_grid$lat)
```

```{r}
for (paths in hycom_path[1]) {
  
hy2 <- hycom_load(paths)
# hy2 <- hycom_load(hycom_path[1])

slice_by <- 200
length_divide <-  2
vel_vect <- slice(hy2$tibble, seq(1, n(), by = slice_by))

xlim <- range(hy2$tibble$lon)
ylim <- range(hy2$tibble$lat)

# ggplot() +
#   geom_tile(data = hycom_grid, aes(x = lon, y = lat, fill = vel_mag)) +
#   # scale_fill_
#   theme_bw() +
#   coord_cartesian(expand = FALSE)
plt <- ggplot() +
  geom_tile(data = hy2$tibble, aes(x = lon, y = lat, fill = vel_mag)) +
    # velocity vectors
  geom_segment(
    data = vel_vect,
    aes(
      x     = lon, 
      xend  = lon + hycom_u/length_divide, 
      y     = lat, 
      yend  = lat + hycom_v/length_divide, 
      color = vel_mag
      ),
    arrow     = arrow(length = unit(0.15/2, "cm")), # adjust arrow size
    linewidth = 0.25,                              # adjust arrow thickness
    na.rm     = TRUE
  ) +
  # scale_fill_
  theme_bw() +
  facet_wrap(~date) +
  scale_fill_viridis_b(breaks = c(0.25, seq(0, 2, by = 0.5))) +
  scale_color_viridis_b(option = "magma", breaks = c(0.25, seq(0, 2, by = 0.5))) +
  coord_cartesian(expand = FALSE, ylim = ylim, xlim = xlim)
print(plt)
}
shelf(camcorder)
camcorder::gg_record(here("data", "plots", "camcorder2"))
camcorder::gg_stop_recording()
camcorder::gg_playback(frame_duration = 0.7)
```

```{r}
hycom_grid %>%
  filter(between(lat, 22.57, 22.58))  %>%
   ggplot() +
  geom_tile(data = hycom_grid, aes(x = lon, y = lat, fill = vel_mag)) +
  # scale_fill_
  theme_bw()
```

