---
title: "HYCOM-TSIS Particle Tracker"
author: "Sebastian Di Geronimo"
date: 2026-07-07
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, #conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  # forcats, lubridate, glue, fs, 
  fs, magrittr, dplyr, here, ggplot2, lubridate,  stringr, ggnewscale, purrr,
  
  # additional
  # sp, raster, ncdf4,
  sf, terra, metR, tidyterra,  geodata
)

# conflicts_prefer(
#   dplyr::filter(), 
#   dplyr::select()
#   )
  
# source aviso functions
source(here("R", "aviso_functions.R"))

# path to executable and path to save data
path_copernicusmarine <- here("copernicusmarine.exe")

path_save <- here("data", "raw", "hycom_particle_tracking")

fs::dir_create(path_save)

plot_out_path <- here("data", "plots", "hycom_particle_tracker", Sys.Date())
plot_out_path <- here("data", "plots", "hycom_particle_tracker", paste0("keys_", Sys.Date()))

fs::dir_create(plot_out_path)
```

## 2.2 Dataset Path

```{r}
# hycom cloud storage
hycom_path <- 
  here(cloud_dir, "data", "raw", "hycom") %>%
  dir_ls() %>%
  dir_ls() %>% 
  dir_ls(regexp = "\\.nc") %T>% {
    print(length(.))
    sample(., 10) %>% print()
  } %>% 
  tibble(file = .) %>% 
  mutate(
    date = str_extract(file, "\\d{4}_\\d{2}_\\d{2}"),
    date = as_date(date)
  ) %T>%
  print()

# filter for cruise dates -7 and +3 days
hycom_filt_path <- 
  hycom_path %>% 
  filter(
    between(
      date, 
      # as_date("2017-10-11") - days(10), 
      # as_date("2017-10-12") + days(5)
      as_date("2017-10-01"), 
      as_date("2017-10-31")
      )
  ) %T>%
  print()
```


```{r load-data}
lat_lon <- 
  c(
    # long
    "xmin" = -83.8125,
    "xmax" = -78.3125,
    #lat
    "ymin" = 22.8125,
    "ymax" = 26.4375
  )

# set date range by month
months <- 
  hycom_filt_path %>% 
  pull(date) %>%
  range() %T>% 
  print()

# months <-
#   seq(
#     lubridate::as_date("2018-01-01"),
#     lubridate::as_date("2018-02-01"),
#     by = "1 month"
#     )

```

URL: <https://data.marine.copernicus.eu/product/WIND_GLO_PHY_L4_MY_012_006/description>
Dataset:         Global Ocean Hourly Reprocessed Sea Surface Wind and Stress
                   from Scatterometer and Model
ID:              cmems_obs-wind_glo_phy_my_l4_0.125deg_PT1H
Resolution:      0.125° × 0.125°
Temporal Extent: 1 Jun 1994 to 18 Feb 2026
Temporal:        Hourly

Variables: 
  Wind
  - eastward_wind - Eastward wind [m/s]
  - northward_wind - Northward wind [m/s]

```{r wind-vel-params}
# set to TRUE if downloading data
download <- FALSE
download <- TRUE

# set to TRUE if want to overwrite previous data
overwrite <- FALSE
# overwrite <- TRUE

# dataset id
dataset_id_wind <- "cmems_obs-wind_glo_phy_my_l4_0.125deg_PT1H"

# select vars
vars <-
  c(
    "eastward_wind",
    "northward_wind"
  )

cat("Date Range: ", format(months[c(1, length(months))], "%B %Y"), sep = "\n- ")

aviso_download(
  months[1], 
  vars, 
  bounds     = lat_lon,
  path_copernicusmarine,
  dataset_id = dataset_id_wind,
  path_save  = path_save,
  download,
  overwrite
)
```
# 3.0 Particle Tracker

## STEP 1: Set Up Functions

```{r}
# function that reads netcdf files and puts them in a workable stack object:
read_mover <- function(ncdfile, xvar = "u", yvar = "v", hycom_time = FALSE) {
  
  u <- terra::rast(ncdfile, subds = xvar)
  v <- terra::rast(ncdfile, subds = yvar)
  
  z_time <- terra::time(u)

  if (hycom_time) {
    z_time <- lubridate::as_date(
      as.numeric(str_extract(names(u), "MT=(.*)", 1)),
      origin = as_date("1900-12-31")
    )
    z_time <- lubridate::as_datetime(z_time)
  }
  
  return(list(u = u, v = v, z = z_time))
}

# function that extrapolates water currents for NA values
extend_data <- function(r, band = 21) {
  terra::focal(
    x         = r,
    # w         = matrix(1, band, band),
    w         = band, # square
    fun       = mean,
    na.rm     = TRUE,
    expand    = TRUE,
    fillvalue = NA,
    na.policy = "only"
  )
}
```

## STEP 2: Load Currents and Wind Data

```{r load-data}
temp <- 
  hycom_filt_path %>% 
  pull(file) %>% 
  read_mover(xvar = "u", yvar = "v", hycom_time = TRUE)

water_u    <- temp$u
water_v    <- temp$v
water_time <- temp$z
rm(temp)

## Read the wind data:
temp       <- 
  here(path_save) %>%
  fs::dir_ls(regexp = dataset_id_wind) %>%
  .[[1]] %>%
  read_mover("northward_wind", "eastward_wind")

wind_u     <- temp$u
wind_v     <- temp$v
wind_time  <- temp$z
rm(temp)
```

```{r plot-vectors}
plot(water_u)
plot(water_v)
plot(wind_u)
plot(wind_v)
```

## STEP 3: Crop and Create Land Mask

```{r country-download}
## Countries for which the shorelines are downloaded and plotted
countries <- c("USA", "Cuba", "Bahamas")

gd_load <- lapply(
  as.list(countries),
  function(x) {
    gadm <- ""

    gadm <- tryCatch(
      geodata::gadm(country = x, level = 0, path = tempdir()),
      error = function(e) NULL
    )

    gadm
  }
)

scale <- 0.05
aspect.ratio <- 1 / cos(mean(lat_lon[3:4]) * pi / 180)
```


```{r country-combine}
land_crop <- land_buff <- list()
base_plt <-
  ggplot() +
  labs(x = NULL, y = NULL) +
  coord_sf(
    xlim = lat_lon[1:2],
    ylim = lat_lon[3:4],
    expand = FALSE
    )  +
  theme_bw()


for (i in seq(length(gd_load))) {
  land_crop[[i]] <- crop(gd_load[[i]], ext(water_u))
  
  # add land to figure
  base_plt <- base_plt + geom_spatvector(data = land_crop[[i]])
  
  # buffer around polygon
  land_buff[[i]] <- buffer(land_crop[[i]], 1000) # meters
  
}

base_plt

# combine all countries to one spatvector
land_union <- purrr::reduce(land_buff, terra::union)

# create raster with the same resolution as currents data
land_mask <- rasterize(
  land_union, 
  rast(ext(water_u), resolution = res(water_u)), 
  field = 1
  ) 
land_mask[is.na(land_mask)] <- 0

plot(land_mask)
```



## STEP 4: Specify Particle Release Details


```{r release-params}
rel_params <- list(
  # time at which the simulated release should start:
  "start"  = as.POSIXct(as_datetime(months[1]), tz = "UTC"),
  
   # time at which the simulated release should end:
  "end"    = as.POSIXct(as_datetime(months[1]) + days(1), tz = "UTC"),
  
  # number of particles to release in total:
  # "num"    = 10,
  "num"    = 100,
  
  # release coordinates:
  # "coords" = c(
  #   -83, # lon
  #   24,  # lat
  #   0    # depth not used in simulation
  #   )
  # "coords" = c( # location of particle start
  #   -81.25, # lon
  #   25.5,  # lat
  #   0    # depth not used in simulation
  #   )
  "coords" = list( # location of particle start
    "lon"   = c(-81.25, -82.6, -83), # lo
    "lat"   = c(25.5, 24.2, 26),  
    "depth" = c(0, 0, 0)   
    )
  ) %T>% 
  print()

# plot start location
base_plt +
  annotate(
    geom = "point",
    # data = rel_params$coords, 
    x = rel_params$coords$lon, y = rel_params$coords$lat,
    color = "red", size = rel(2)
  )

```

## STEP 5: Specify Simulation Details

```{r simulation-param}
# time at which the simulation should start and end:
sim_param <- list(
  # time at which the simulation should start:
  "start"    = as.POSIXct(as_datetime(months[1]), tz = "UTC"),
  
  # time at which the simulation should stop:
  "end"      = as.POSIXct(as_datetime(months[2]), tz = "UTC"), # end
  # "end"      = as.POSIXct(as_datetime(months[1]) + days(2), tz = "UTC"), # end
  
  # time step of simulation in seconds:
  # "step"     = 60 * 60 / 4, # seconds per step 0.25 hr
  "step"     = 60 * 60 * 6, # seconds per step 6 hr
  # "step"     = 60 * 60 * 3, # seconds per step 3 hr
  # "step"     = 60 * 60, # seconds per step 1 hr
  
  # value indicating every n-th time step that should be exported:
  "plt_step" = 3, 
  
  # set direction of simulation
  "direct" = "fwd" # forward
  # "direct" = "bck" # backward
) %T>% 
  print()

# when simulation occurs
# create a time series based on the start and end time and the time step:
simulation_time <- seq(
  sim_param[["start"]], 
  sim_param[["end"]] - sim_param[["step"]], 
  sim_param[["step"]]
  ) %T>%
  print()

```

## STEP 6: Set How Many Paricles and Where

```{r particle-setup}
# for each time step determine how many particles are released:
n_particles           <- rep(NA, length(simulation_time))
# n_particles[simulation_time <= release_start] <- 0
# n_particles[simulation_time >= release_end]   <- release_n
n_particles[simulation_time <= rel_params[["start"]]] <- 0
n_particles[simulation_time >= rel_params[["end"]]]   <- rel_params[["num"]]  * length(rel_params$coords$lon)

n_particles[is.na(n_particles)] <-
  approx(
    c(0, rel_params$num),
    n = 2 + sum(is.na(n_particles))
  )$y[c(-1, - (2 + sum(is.na(n_particles))))]

n_particles           <- round(n_particles)

# create a SpatVector for the particles:
particles <-
  data.frame(
    pos_x = rep(rel_params$coords$lon,  rel_params$num),
    pos_y = rep(rel_params$coords$lat,  rel_params$num),
    pos_z = rep(rel_params$coords$depth,  rel_params$num),
    id = rep(letters[seq(length(rel_params$coords$lon))], rel_params$num)
  ) %>%
  terra::vect(
    geom = c("pos_x", "pos_y"),
    crs  = "EPSG:4326" # WGS84
  )

# projection to be used to calculate particle displacement in meters:
part_proj <- terra::crs("EPSG:6442") # West Florida

# drift factor indicates with which fraction particles are displaced by the wind speed:
drift_factor    <- 0.03

# the diffusion coefficient, used in random walk simulation
diffusion_coeff <- 1e5
```

## STEP 7: Run Particle Releae Simulation

TODO: add coastline boundary
TODO: add in if NA extract then replace with previous value

```{r}
# create a list in which results should be stored:
part_results          <- list()
previous_position     <- particles

# set direction to multiple vectors by
signs <- switch(
  sim_param$direct,
  "fwd" = 1,
  "bck" = -1
)

# set figure numbers based on forward or backward
out_list <- switch(
  sim_param$direct,
  "fwd" = 0,
  "bck" = length(which(((seq(length(simulation_time)))- 1) %% 3 == 0)) + 1
)

# set direction of time steps
steps <- switch(
  sim_param$direct,
  "fwd" = seq(length(simulation_time)),
  "bck" = rev(seq(length(simulation_time)))
)

## start looping the simulation times
# for (i in steps) {
for (i in steps[1:16]) {
  # message(paste(i, "of", length(simulation_time)))
  print(paste(i, "of", length(simulation_time)))
  
  # Get the water currents that are closest to the current simulation time
  tdif   <- abs(water_time - simulation_time[i])
  tsel_c <- which(tdif == min(tdif))[[1]]
  twat_u <- extend_data(water_u[[tsel_c]]) * signs
  twat_v <- extend_data(water_v[[tsel_c]]) * signs
  
  
  tdif   <- abs(wind_time - simulation_time[i])
  tsel_w <- which(tdif == min(tdif))[[1]]
  twin_u <- extend_data(wind_u[[tsel_w]]) * signs
  twin_v <- extend_data(wind_v[[tsel_w]]) * signs

  # Extract last known particle locations and transform coordinates into meters:
  # 1. transforms decimal degree to meters as matrix x, y coords
  # 2. extract coordinates
  part_coords <- prev_coords <- 
    terra::project(previous_position, part_proj) %>% 
    terra::crds()

  
  # displace particles with *water currents*
  part_coords[seq(0, n_particles[i]), ] <- 
    part_coords[seq(0, n_particles[i]), ] +
    sim_param[["step"]] * # integrate m/s to m i.e m/s * seconds
    # extract u and v components where particle is nearest
    # bilinear = 4 nearest cells
    cbind(
      "x" = terra::extract(twat_u, previous_position, method = "bilinear", ID = FALSE)[, 1],
      "y" = terra::extract(twat_v, previous_position, method = "bilinear", ID = FALSE)[, 1]
      )[seq(0, n_particles[i]), ]
  
  if (any(is.nan(part_coords))) {
    bad_rows <- unique(which(arr.ind = TRUE, is.nan(part_coords))[, 1])
    part_coords[bad_rows, ] <- prev_coords[bad_rows,]
    prev_coords <- part_coords
  }
  
  ## displace particles with *wind*
  part_coords[seq(0, n_particles[i]),] <- 
    part_coords[seq(0, n_particles[i]),] +
    sim_param[["step"]] * drift_factor * 
    cbind(
      "x" = terra::extract(twin_u, previous_position, method = "bilinear", ID = FALSE)[, 1],
      "y" = terra::extract(twin_v, previous_position, method = "bilinear", ID = FALSE)[, 1]
      )[seq(0, n_particles[i]), ]
  

  if (any(is.nan(part_coords))) {
    bad_rows <- unique(which(arr.ind = TRUE, is.nan(part_coords))[, 1])
    part_coords[bad_rows, ] <- prev_coords[bad_rows,]
   }
  
  # displace with *random walk* (diffusion):
  part_coords[0:n_particles[i],] <- 
    part_coords[0:n_particles[i], ] +
    cbind(
      "x" = runif(n_particles[i], -1, 1),
      "y" = runif(n_particles[i], -1, 1)
      ) * sqrt(diffusion_coeff * sim_param[["step"]] * 6 / 1e4)

  
  # turn particle information into a spatvector object:
  part_spv <-
    as.data.frame(part_coords) %>%
    vect(geom = c("x", "y"), crs = part_proj) %>%
    project(crs(previous_position))

  ## Coastline blocking with slide-along
  # Previous positions in projected meter space
  prev_proj <- terra::project(previous_position, part_proj)
  prev_xy   <- terra::crds(prev_proj)
  
  # New candidate positions in the same projected space
  new_proj  <- terra::project(part_spv, part_proj)
  new_xy    <- terra::crds(new_proj)
  
  # ---- Check land mask at new positions ---- #
  land_vals <- terra::extract(
    land_mask, 
    new_proj, 
    method = "simple", 
    ID = FALSE
    )
  ok <- (land_vals == 0)  # 0 = water, 1 = land
  
  # 1 = land, 0 = water, NA = outside domain
  is_land   <- !is.na(land_vals) & (land_vals == 1)
  in_domain <- !is.na(land_vals)
  
  ok <- !is_land & in_domain  # keep only if water AND inside domain
  beached_or_out <- !ok       # mark for removal / beaching
  
  if (any(!ok)) {
    # For particles that hit land, slide them back along the segment
    idx <- which(!ok & in_domain)
    
    if (length(idx) != 0) {
      for (j in idx) {
         # old (water)
        x0 <- prev_xy[j, 1] 
        y0 <- prev_xy[j, 2]
        
        # new (land)
        x1 <- new_xy[j, 1] 
        y1 <- new_xy[j, 2]    
    
        # Binary search along the segment for the last water point
        a <- 0
        b <- 1
        
        for (k in 1:10) {
          t  <- (a + b) / 2
          xt <- x0 + t * (x1 - x0)
          yt <- y0 + t * (y1 - y0)
    
          tmp_spv <- vect(
            data.frame(x = xt, y = yt),
            geom = c("x", "y"),
            crs  = part_proj
            ) %>%
            project(crs(land_mask))
    
          val <- terra::extract(land_mask, tmp_spv, method = "simple", ID = FALSE)
          if (is.na(val)) val <- 0  # treat NA as water if needed
    
          if (val == 0) {
            a <- t  # still in water, go further
          } else {
            b <- t  # in land, go back
          }
        }
    
        # Use point just inside water
        t_final <- a
        new_xy[j, 1] <- x0 + t_final * (x1 - x0)
        new_xy[j, 2] <- y0 + t_final * (y1 - y0)
      }
  
    part_spv <- vect(
      as.data.frame(new_xy),
      geom = c("x", "y"),
      crs  = part_proj
    ) %>%
      project(crs(previous_position))
    }
  }

  # store the new locations in the results list:
  previous_position <- part_spv
  
  # ---- plot ---- #
  
  # for each n-th simulation step store the results (where n = ["plt_step"]]):
  if ((i - 1) %% sim_param[["plt_step"]] == 0) {
    
    out_list <- out_list + signs
    
    time_step <- paste(
      format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
      attr(as.POSIXlt(simulation_time[i]),"tzone")
      )
    
    # add an indicator of the release state of the particles:
    part_spv$released <- rep(FALSE, rel_params$num)
    
    # part_coords$released[0:n_particles[i]] <- TRUE
    part_spv$released[0:n_particles[i]] <- TRUE
    
    part_spv$part_id <- particles$id
    part_spv$part_num <- seq(length(particles$id))
    
    
    # part_results[[out_list]] <- part_coords
    part_results[[out_list]] <- part_spv
    names(part_results)[out_list] <- time_step
    
    # extract water currents x, y and magnitude
    wat_arrow <- 
      dplyr::left_join(
        as.data.frame(water_u[[tsel_c]], xy = TRUE),
        as.data.frame(water_v[[tsel_c]], xy = TRUE)
      ) %>%
      rename(
        "u" = contains("u_"),
        "v" = contains("v_")
      ) %>%
      mutate(
        mag  = sqrt(u^2 + v^2),
        time = time_step
        )

    # extract wind speed x, y and magnitude
    win_arrow <- 
      dplyr::left_join(
        as.data.frame(wind_u[[tsel_w]], xy = TRUE),
        as.data.frame(wind_v[[tsel_w]], xy = TRUE) 
      ) %>%
      rename(
        "north" = contains("north"),
        "east"  = contains("east")
      ) %>%
      mutate(
        mag   = sqrt(north^2 + east^2),
        north = north * scale,
        east  = east * scale,
        time  = time_step
        )
    
    # create image file
    png(
      here(plot_out_path, sprintf("out%03d.png", out_list)),
      width  = 2000,
      height = 1500,
      res    = 300,
      type   = "cairo"
      ) 
  
  
    print( 
      base_plt +
        
      # wind
      metR::geom_arrow(
        data    = win_arrow,
        mapping = aes(x = x, dx = north, y = y, dy = east, color = mag),
        alpha   = 0.75,
        skip    = 1,
        size    = 0.5, # adjust arrow thickness
        na.rm   = TRUE
        ) +
      scale_color_viridis_c(
        name   = "Wind Speed", 
        breaks = seq(0, 20, 5),
        limits = c(0, 20),
        oob    = scales::oob_squish
        ) +  
        
      ggnewscale::new_scale_color() +
      
      # water  
      metR::geom_arrow(
        data    = wat_arrow,
        mapping = aes(x = x, dx = u, y = y, dy = v, color = mag),
        skip    = 10,
        size    = 0.5, # adjust arrow thickness
        na.rm   = TRUE
        ) +
      scale_color_cross_blended_c(
        name   = "Current", 
        breaks = seq(0, 3, 1),
        limits = c(0, 3),
        oob    = scales::oob_squish,
        )  +
      
      ggnewscale::new_scale_color() +
      
      # particles  
      tidyterra::geom_spatvector(
        data  = part_results[[out_list]],
        # color = "red",
        aes(color = part_id)
        ) +
      
      scale_color_discrete(name = "ID") +  
        
      labs(
        title = paste(
          format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
          attr(as.POSIXlt(simulation_time[i]),"tzone"))
      ) +
   
      coord_sf(
        xlim = plot.xlim,
        ylim = plot.ylim,
        expand = FALSE
        ) 
    )

    dev.off()
    
    # out_list <- out_list + signs
  }
  
  
}


```
## STEP 8: Convert Figures into Video using `ffmpeg`

```{r}
system2(
  here(ffmpeg_path),
  paste(
    "-framerate 1 -i",
    here(plot_out_path, "out%03d.png"),
    "-c:v libx264 -pix_fmt yuv420p",
    here(dirname(plot_out_path), paste0(
      "particle3_n_", rel_params$num, "_", format(Sys.Date(), "%Y_%m_%d"), ".mp4"
      )
    )
  )
)



```

## STEP 9: Save Particle Trajectories

```{r}
# save the resulting particle information:
all_results <- 
  purrr::map(part_results, \(.x) sf::st_as_sf(.x)) %>% 
  purrr::list_rbind(names_to = "time") %>% 
  sf::st_as_sf() %>% 
  mutate(
    time = as_datetime(time),
    lon  = st_coordinates(geometry)[,1],
    lat  = st_coordinates(geometry)[,2],
    )

readr::write_csv(all_results, here(dirname(plot_out_path), "test.csv"))
```


```{r}
ggplot() +
    # base_plt +
    geom_line(
      data = test_p,
      aes(
        x = lon, 
        y = lat,
        color = part_id,
        group = part_num),
      linetype = "dashed",
      alpha = 0.5
      ) +
  geom_sf(
    data = test_p,
  aes(color = part_id)
  ) +
    theme()

plotly::ggplotly(
    ggplot() +
    # base_plt +
    geom_line(
      data = test_p,
      aes(
        x = st_coordinates(geometry)[,1], 
        y = st_coordinates(geometry)[,2],
        color = part_id,
        group = part_num),
      linetype = "dashed",
      alpha = 0.5
      ) +
  geom_sf(
    data = test_p,
  aes(color = part_id, group = part_num, time = time)
  ) +
    theme()
)

```


# ERA5 Wind Data from csv
```{r}
wind_dir <- rstudioapi::selectDirectory()

wind_path <- dir_ls(wind_dir) %>% 
  print()
```

```{r}
era_wind <- 
  wind_path[1] %>% 
  readr::read_csv(show_col_types = FALSE)

era_wind %>% 
  filter(
    date_utc == "2022-07-30" & hour_utc == 5
  ) %>%
  ggplot() +
  geom_raster(
    aes(x = longitude, y = latitude, fill = wind_speed_m_s)
  ) +
  scale_fill_viridis_c()

era_wind %>% 
  filter(
    date_utc == "2022-07-30" & hour_utc == 5
  ) %>%
  mutate(
    # .keep = "used",
    # mag = sqrt(u10_m_s^2 + v10_m_s^2),
    # diff = mag - wind_speed_m_s
  ) %>% 
  ggplot() +
 # wind
      metR::geom_vector(
        # data    = win_arrow,
        mapping = aes(
          x = longitude, 
          y = latitude, 
          dx = u10_m_s * scale * 2, 
          dy = v10_m_s * scale * 2, 
          color = wind_speed_m_s
          ),
        alpha   = 0.75,
        skip    = 1,
        size    = 0.5, # adjust arrow thickness
        na.rm   = TRUE
        ) +
      scale_color_viridis_c(
        name   = "Wind Speed", 
        breaks = seq(0, 20, 5),
        limits = c(0, 20),
        oob    = scales::oob_squish
        ) +
  scale_mag()
```


```{r}
era_wind %>% 
  filter(
    date_utc == "2022-07-30" & hour_utc == 5
  ) %>%
  mutate(
    # .keep = "used",
    # mag = sqrt(u10_m_s^2 + v10_m_s^2),
    # diff = mag - wind_speed_m_s
    angle = Angle(u10_m_s, v10_m_s)
  ) %>% 
  slice(c(70 )) %>%
  ggplot() +
 # wind
      metR::geom_vector(
        # data    = win_arrow,
        mapping = aes(
          x = longitude, 
          y = latitude,    
          mag = wind_speed_m_s*10,
          angle =  angle,
          color = wind_speed_m_s
          ),
        alpha   = 0.75,
        skip    = 0,
        size    = 0.5, # adjust arrow thickness
        na.rm   = TRUE
        ) +
      scale_color_viridis_c(
        name   = "Wind Speed", 
        breaks = seq(0, 20, 5),
        limits = c(0, 20),
        oob    = scales::oob_squish
        ) 


```


```{r}
ggplot() +
  # annotate(
  #   "segment",
  #   x = 0,
  #   y = 0,
  #   xend = -3.9649353,
  #   yend = 0.223754883
  # ) +
  geom_arrow(
    # data = tibble(
    #   x = 0,
    #   y = 0,
    #   angle = 273.2300,
    #   mag = 3.971244,
    #   color = 3.971244,
    #   dx = -3.9649353,
    #   dy = 0.223754883
    # ),
    data = era_wind %>% 
  filter(
    date_utc == "2022-07-30" & hour_utc == 5
  ) %>% slice(1),
    aes(
      x = longitude,
      y = latitude,
      dx = u10_m_s,
      dy = v10_m_s,
      # angle = Angle(u10_m_s, v10_m_s),
      # mag = Mag(u10_m_s, v10_m_s)
    ),
    color = "red",
    pivot = 0,
    direction = "cw"
  ) +
  scale_mag() +
  # geom_hline(yintercept = 0, linetype = "dashed") +
  # geom_vline(xintercept = 0, linetype = "dashed") +
  # # coord_cartesian(xlim = c(-7, 7), ylim = c(-7, 7)) +
  theme()
  
```

