---
title: "2D Particle Tracking Test"
author: "Sebastian Di Geronimo"
date: 2026-06-26
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

From blog:
<https://r-coders-anonymous.blogspot.com/2016/07/2d-particle-tracking-in-r.html>

This uses old packages that will probably need updating:

sp
raster
ncdf4
ncdf.tools - not available
rgeos - not available
maptools - not available


## Download Data from `Copernicusmarine` App

- using `copernicusmarine.exe` File

From: <https://help.marine.copernicus.eu/en/articles/10750437-copernicus-marine-toolbox-executable-no-installation>

I recommend moving the file to this directory where this project is stored

To start, open `cmd` or `PowerShell` from explorer. `cd` to location where `.exe`
is located. Then run `copernicusmarine login` and enter "Username" and
"Password". Once completed, you should be able to run the download

Download Command example:

copernicusmarine subset
--dataset-id cmems_mod_glo_phy_my_0.083deg_P1D-m
   --variable vo
   --variable uo
   --start-datetime 2018-01-01T00:00:00
   --end-datetime 2018-02-01T00:00:00
   --minimum-longitude -3.26
   --maximum-longitude 9.6
   --minimum-latitude 50.62
   --maximum-latitude 62.67
   --minimum-depth 0.49402499198913574 <----- for water depth
   --maximum-depth 0.49402499198913574 <----- for water depth


# 2.0 ---- Setup ----

## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, #conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  # forcats, lubridate, glue, fs, 
  fs, magrittr, dplyr, here, ggplot2, lubridate,  stringr, ggnewscale, 
  
  # additional
  # sp, raster, ncdf4,
  terra, metR, tidyterra,  geodata
)

# conflicts_prefer(
#   dplyr::filter(), 
#   dplyr::select()
#   )

# source aviso functions
source(here("R", "aviso_functions.R"))

# path to executable and path to save data
path_copernicusmarine <- here("copernicusmarine.exe")

path_save <- here("data", "raw", "particle_tracking")

fs::dir_create(path_save)

plot_out_path <- here("data", "plots", "particle_tracker", Sys.Date())

fs::dir_create(plot_out_path)
```



## 2.2 Download Data

URL: <https://data.marine.copernicus.eu/product/GLOBAL_MULTIYEAR_PHY_001_030/description>
Dataset:         Global Ocean Physics Reanalysis
ID:              cmems_mod_glo_phy_my_0.083deg_P1D-m
Resolution:      0.083° × 0.083°
Temporal Extent: 1 Jan 1993 to 26 May 2026
Temporal:        Daily

Variables: 
  Water current:
  - vo - eastward sea water velocity [m/s]
  - uo - northward sea water velocity [m/s]

URL: <https://data.marine.copernicus.eu/product/WIND_GLO_PHY_L4_MY_012_006/description>
Dataset:         Global Ocean Hourly Reprocessed Sea Surface Wind and Stress
                   from Scatterometer and Model
ID:              cmems_obs-wind_glo_phy_my_l4_0.125deg_PT1H
Resolution:      0.125° × 0.125°
Temporal Extent: 1 Jun 1994 to 18 Feb 2026
Temporal:        Hourly

Variables: 
  Wind
  - eastward_wind - Eastward wind [m/s]
  - northward_wind - Northward wind [m/s]


NOTE: When saving data from copernicus, the data can be stacked with multiple
time points so should breakup the date ranges as storing as 1 month at a time 
chunks

Spatial:
N: 62.67 ymax
W: -3.26 xmin
S: 50.62 ymin
E: 9.60 xmax

```{r base-params}
# set date range by month
months <-
  seq(
    lubridate::as_date("2018-01-01"),
    lubridate::as_date("2018-02-01"),
    by = "1 month"
    )

# set spatial extent
part2d_bbox <-
  c( 
    xmin = -3.26,
    xmax = 9.60,
    ymin = 50.62,
    ymax = 62.67
  )

# set to TRUE if downloading data
download <- FALSE
# download <- TRUE

# set to TRUE if want to overwrite previous data
overwrite <- FALSE
# overwrite <- TRUE
```


```{r water-current-params}
# dataset id
dataset_id_phy <- "cmems_mod_glo_phy_my_0.083deg_P1D-m"

# select vars
vars <-
  c(
    "vo",
    "uo"
  )

add_params <-
  c(
    "--minimum-depth 0.49402499198913574", 
    "--maximum-depth 0.49402499198913574"
    )

cat("Date Range: ", format(months[c(1, length(months))], "%B %Y"), sep = "\n- ")

aviso_download(
  months[1], 
  vars, 
  bounds      = part2d_bbox,
  path_copernicusmarine,
  dataset_id  = dataset_id_phy,
  path_save   = path_save,
  download,
  overwrite,
  other_param = add_params
)
```


```{r wind-vel-params}
# dataset id
dataset_id_wind <- "cmems_obs-wind_glo_phy_my_l4_0.125deg_PT1H"

# select vars
vars <-
  c(
    "eastward_wind",
    "northward_wind"
  )

cat("Date Range: ", format(months[c(1, length(months))], "%B %Y"), sep = "\n- ")

aviso_download(
  months[1], 
  vars, 
  bounds     = part2d_bbox,
  path_copernicusmarine,
  dataset_id = dataset_id_wind,
  path_save  = path_save,
  download,
  overwrite
)
```

# 3.0 2D Particle Tracking using `terra`  


## STEP 1: load the water currents and wind data


```{r func}
shelf(terra)

## function that reads netcdf files and puts them in a workable stack object:
read_mover <- function(ncdfile, xvar = "u", yvar = "v") {
  
  u <- terra::rast(ncdfile, subds = xvar)
  v <- terra::rast(ncdfile, subds = yvar)
  
  z_time <- terra::time(u)

  return(list(u = u, v = v, z = z_time))
}

## Countries for which the shorelines are downloaded and plotted
countries <- c("NLD", "GBR", "FRA")

gd_load <- lapply(
  as.list(countries),
  function(x) {
    gadm <- ""
    lev <- 0
    if (x == "NLD") {
      lev <- 1
    } else {
      lev <- 0
    }

    gadm <- tryCatch(
      geodata::gadm(country = x, level = lev, path = tempdir()),
      error = function(e) NULL
    )

    if (inherits(gadm, "SpatVector") && lev == 1 && !is.null(gadm)) {
      gadm <- gadm[gadm$ENGTYPE_1 != "Water body", ]
      gadm <- terra::aggregate(gadm, by = "GID_1")
    }

    gadm
  }
)

base_plt <- ggplot()
   
for (i in seq(length(gd_load))) {
  base_plt <- base_plt + geom_spatvector(data = gd_load[[i]]) 
}

# function that extrapolates water currents for NA values
extend_data <- function(r, band = 21) {
  terra::focal(
    x         = r,
    # w         = matrix(1, band, band),
    w         = band, # square
    fun       = mean,
    na.rm     = TRUE,
    expand    = TRUE,
    fillvalue = NA,
    na.policy = "only"
  )
}
```


```{r load-data}
## Read the water current data:
temp       <-
  here(path_save) %>%
  fs::dir_ls(regexp = dataset_id_phy) %>%
  .[[3]] %>%
  read_mover("uo", "vo")
water_u    <- temp$u
water_v    <- temp$v
water_time <- temp$z

set.names(water_u, stringr::str_remove(names(water_u), "=\\d.\\d+"))
set.names(water_v, stringr::str_remove(names(water_v), "=\\d.\\d+"))
rm(temp)

## Read the wind data:
temp       <- 
  here(path_save) %>%
  fs::dir_ls(regexp = dataset_id_wind) %>%
  .[[3]] %>%
  read_mover("northward_wind", "eastward_wind")
wind_u     <- temp$u
wind_v     <- temp$v
wind_time  <- temp$z
rm(temp)
```


```{r plot-vectors}
plot(water_u)
plot(water_v)
plot(wind_u)
plot(wind_v)
```

## STEP 2: specify particle release details


```{r}
# time at which the simulated release should start:
release_start         <- as.POSIXct("2018-01-02 08:00", tz = "UTC")

# time at which the simulated release should end:
release_end           <- as.POSIXct("2018-01-04 08:00", tz = "UTC")

# number of particles to release in total:
# release_n             <- 5000
release_n             <- 10

# release lon-coordinate:
release_x             <- 3.7

# release lat-coordinate:
release_y             <- 52

# release depth (not used yet in simulation)
release_z             <- 0
```


## STEP 3: specify simulation details


```{r}
# time at which the simulation should start:
simulation_start_time <- as.POSIXct("2018-01-02 00:00", tz = "UTC")

# time step of simulation in seconds:
# simulation_time_step  <- 60 * 60 / 4 # 15 min
simulation_time_step  <- 60 * 60 * 6 # 6 hours

# integer value indicating every n-th time step that should be exported:
simulation_out_step   <- 2

# time at which the simulation should stop
simulation_end_time   <- as.POSIXct("2018-01-15 08:00", tz = "UTC")

# create a time series based on the start and end time and the time step:
simulation_time       <- seq(
  simulation_start_time, 
  simulation_end_time - simulation_time_step, 
  simulation_time_step
  )
```


```{r}
# for each time step determine how many particles are released:
n_particles           <- rep(NA, length(simulation_time))
n_particles[simulation_time <= release_start] <- 0
n_particles[simulation_time >= release_end]   <- release_n
n_particles[is.na(n_particles)]               <- approx(
  c(0, release_n), 
  n = 2 + sum(is.na(n_particles))
  )$y[c(-1, -(2 + sum(is.na(n_particles))))]

n_particles           <- round(n_particles)

# create a SpatialPointsDataFrame for the particles:
particles <-
  data.frame(
    pos_x = rep(release_x, release_n),
    pos_y = rep(release_y, release_n),
    pos_z = rep(release_z, release_n)
  ) %>%
  terra::vect(
    geom = c("pos_x", "pos_y"),
    crs  = "EPSG:4326" # WGS84
  )

# projection to be used to calculate particle displacement in meters:
part_proj             <- terra::crs("EPSG:32631")

# drift factor indicates with which fraction particles are displaced by the wind speed:
drift_factor    <- 0.03

# the diffusion coefficient, used in random walk simulation
diffusion_coeff <- 1e5
```

## STEP 4: loop all simulation time steps

```{r}
# plot(rast(water_u, subds  = tsel_c))
plot(water_u[[tsel_c]])
plot(extend_data(water_u[[tsel_c]]))

plot(wind_v[[tsel_w]])
plot(twin_v)
```


```{r}
# create a list in which results should be stored:
part_results          <- list()
previous_position     <- particles

out_list <- 1
# i = 1 # rm later

## start looping the simulation times
for (i in 1:length(simulation_time)) {
  print(i)
  ## Get the water currents that are closest to the current simulation time
  tdif   <- abs(water_time - simulation_time[i])
  tsel_c <- which(tdif == min(tdif))[[1]]
  twat_u <- extend_data(water_u[[tsel_c]])
  twat_v <- extend_data(water_v[[tsel_c]])
  
  
  tdif   <- abs(wind_time - simulation_time[i])
  tsel_w <- which(tdif == min(tdif))[[1]]
  twin_u <- extend_data(wind_u[[tsel_w]])
  twin_v <- extend_data(wind_v[[tsel_w]])

  ## Extract last known particle locations and transform coordinates into meters:
  # 1. transforms decimal degree to meters as matrix x, y coords
  # 2. extract coordinates
  part_coords <- 
    terra::project(previous_position, part_proj) %>% 
    terra::crds()

  ## displace particles with *water_currents*
  part_coords[seq(0, n_particles[i]), ] <- 
    part_coords[seq(0, n_particles[i]), ] +
    simulation_time_step * # integrate m/s to m i.e m/s * seconds
    # extract u and v components where particle is nearest
    # bilinear = 4 nearest cells
    cbind(
      "x" = terra::extract(twat_u, previous_position, method = "bilinear", ID = FALSE)[, 1],
      "y" = terra::extract(twat_v, previous_position, method = "bilinear", ID = FALSE)[, 1]
      )[seq(0, n_particles[i]), ]
  
  ## displace particles with *wind*
  part_coords[seq(0, n_particles[i]),] <- 
    part_coords[seq(0, n_particles[i]),] +
    simulation_time_step * drift_factor * 
    cbind(
      "x" = terra::extract(twin_u, previous_position, method = "bilinear", ID = FALSE)[, 1],
      "y" = terra::extract(twin_v, previous_position, method = "bilinear", ID = FALSE)[, 1]
      )[seq(0, n_particles[i]), ]
  
  
  ## displace with *random walk* (diffusion):
  part_coords[0:n_particles[i],] <- 
    part_coords[0:n_particles[i], ] +
    cbind(
      "x" = runif(n_particles[i], -1, 1),
      "y" = runif(n_particles[i], -1, 1)
      ) * sqrt(diffusion_coeff * simulation_time_step * 6 / 1e4)

  ## turn particle information into a spatvector object:
  part_coords <- 
    as.data.frame(part_coords) %>%
    vect(
      geom = c("x", "y"),
      crs  = part_proj
    ) %>%
    project(crs(previous_position))

  ## store the new locations in the results list:
  previous_position     <- part_coords

  ## for each n-th simulation step store the results (where n = simulation_out_step):
  if ((i - 1) %% simulation_out_step == 0) {
    
    # add an indicator of the release state of the particles:
    part_coords$released <- rep(FALSE, release_n)
    
    part_coords$released[0:n_particles[i]] <- TRUE
    
    part_results[[out_list]] <- part_coords
    names(part_results)[out_list] <- paste(
      format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
      attr(as.POSIXlt(simulation_time[i]),"tzone")
      )
    
    
    # plot limits by adding 2 degrees 
    # add_deg <- 0.5
    # plot.lims <- apply(crds(part_coords), 2, mean)
    # plot.xlim <- c(
    #   "xmin" = plot.lims[[1]] - add_deg, 
    #   "xmax" = plot.lims[[1]] + add_deg
    #   )
    # plot.ylim <- c(
    #   "ymin" = plot.lims[[2]] - add_deg, 
    #   "ymax" = plot.lims[[2]] + add_deg
    #   )
    plot.xlim <- c(
      "xmin" = 2,
      "xmax" = 5
      )
    plot.ylim <- c(
      "ymin" = 51, 
      "ymax" = 56
      )
    
    scale <- 0.05
    aspect.ratio <- 1 / cos(mean(plot.ylim) * pi / 180)
    
    wat_arrow <- dplyr::left_join(
      as.data.frame(water_u[[tsel_c]], xy = TRUE),
      as.data.frame(water_v[[tsel_c]], xy = TRUE)
      ) %>%
      rename(
        "u" = 3,
        "v" = 4
      ) %>%
      mutate(mag = sqrt(u^2 + v^2))

    win_arrow <- dplyr::left_join(
      as.data.frame(wind_u[[tsel_w]], xy = TRUE),
      as.data.frame(wind_v[[tsel_w]], xy = TRUE) 
      ) %>%
      rename(
        "north" = 3,
        "east" = 4
      ) %>%
      mutate(
        mag = sqrt(north^2 + east^2),
        north = north * scale,
        east  = east * scale
        )
    
    png(
      here(plot_out_path, sprintf("out%03d.png", out_list)),
      width = 2000,
      height = 1500,
      res    = 300,
      type   = "cairo"
      ) 
  
  
    print( 
      base_plt +
        
      # wind
      metR::geom_arrow(
        data = win_arrow,
        mapping = aes(x = x, dx = north, y = y, dy = east, color = mag),
        alpha = 0.75,
        skip = 1,
        size = 0.5,                              # adjust arrow thickness
        na.rm = TRUE
        ) +
      scale_color_viridis_c(
        name   = "Wind Speed", 
        breaks = seq(0, 20, 5),
        limits = c(0, 20),
        oob    = scales::oob_squish
        ) +  
        
      ggnewscale::new_scale_color() +
      
      # water  
      metR::geom_arrow(
        data    = wat_arrow,
        mapping = aes(x = x, dx = u, y = y, dy = v, color = mag),
        skip    = 1,
        size    = 0.5,                              # adjust arrow thickness
        na.rm   = TRUE
        ) +
      scale_color_cross_blended_c(
        name   = "Current", 
        breaks = seq(0, 3, 1),
        limits = c(0, 3),
        oob    = scales::oob_squish,
        )  +
      
      # particles  
      tidyterra::geom_spatvector(
        data = part_results[[out_list]],
        color = "red"
        ) +
      labs(
        title = paste(
          format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
          attr(as.POSIXlt(simulation_time[i]),"tzone")),
        x = NULL,
        y = NULL
      ) +
   
      coord_sf(
        xlim = plot.xlim,
        ylim = plot.ylim
        ) +
      theme_bw()
    )

    dev.off()
    
    out_list <- out_list + 1
  }
}

## save the resulting particle information:
# save(part_results, file = "output/out.RData", compress = TRUE)
```



```{r}
# for (i in seq_along(simulation_time)) {
#   print(i)
# 
#   tdif   <- abs(as.numeric(water_time) - as.numeric(simulation_time[i]))
#   tsel_c <- which(tdif == min(tdif))[[1]]
# 
#   # twat_u <- extend_data(terra::rast(water_u, tsel_c))
#   twat_u <- extend_data(terra::subset(water_u, tsel_c))
#   # twat_v <- extend_data(terra::rast(water_v, tsel_c))
#   twat_v <- extend_data(terra::subset(water_v, tsel_c))
# 
#   tdif   <- abs(as.numeric(wind_time) - as.numeric(simulation_time[i]))
#   tsel_w <- which(tdif == min(tdif))[[1]]
# 
#   # twin_u <- extend_data(terra::rast(wind_u, tsel_w))
#   # twin_v <- extend_data(terra::rast(wind_v, tsel_w))
#   twin_u <- extend_data(terra::subset(wind_u, tsel_w))
#   twin_v <- extend_data(terra::subset(wind_v, tsel_w))
# 
#   # issues start here
#   stop()
#   
#   pproj <- terra::project(previous_position, part_proj)
#   part_coords <- terra::crds(pproj)
#   
#   # part_coords <- terra::crds(terra::project(previous_position, part_proj))
# 
#   uv_wat <- cbind(
#     terra::extract(twat_u, pproj, method = "bilinear", ID = FALSE),
#     terra::extract(twat_v, pproj, method = "bilinear", ID = FALSE)
#   )
# 
#   part_coords[seq_along(n_particles[i]), ] <-
#     part_coords[seq_along(n_particles[i]), ] +
#     simulation_time_step * uv_wat[seq_along(n_particles[i]), ]
# 
#   uv_win <- cbind(
#     terra::extract(twin_u, pproj, method = "bilinear", ID = FALSE),
#     terra::extract(twin_v, pproj, method = "bilinear", ID = FALSE)
#   )
# 
#   part_coords[1:n_particles[i], ] <-
#     part_coords[1:n_particles[i], ] +
#     simulation_time_step * drift_factor * uv_win[1:n_particles[i], ]
# 
#   part_coords[1:n_particles[i], ] <-
#     part_coords[1:n_particles[i], ] +
#     cbind(
#       runif(n_particles[i], -1, 1),
#       runif(n_particles[i], -1, 1)
#     ) * sqrt(diffusion_coeff * simulation_time_step * 6 / 1e4)
# 
#   part_coords <- as.data.frame(part_coords)
#   particles <- terra::vect(
#     part_coords,
#     geom = c("pos_x", "pos_y"),
#     crs = part_proj
#   )
# 
#   previous_position <- terra::project(particles, terra::crs(previous_position))
# 
#   if ((i - 1) %% 4 == 0) {
#     part_results[[(i - 1)/4 + 1]] <- previous_position
#     names(part_results)[(i - 1)/4 + 1] <- paste(
#       format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
#       attr(as.POSIXlt(simulation_time[i]), "tzone")
#     )
# 
#     png(sprintf("output\\out%03d.png", (i - 1)/4), 640, 480, res = 75, type = "cairo")
#     plot_ylim <- c(51.9, 53)
# 
#     plot(
#       part_results[[(i - 1)/4 + 1]],
#       pch = ".",
#       col = "black",
#       main = paste(
#         format(simulation_time[i], "%Y-%m-%d %H:%M:%S"),
#         attr(as.POSIXlt(simulation_time[i]), "tzone")
#       )
#     )
# 
#     wat_arrow <- as.data.frame(terra::crds(terra::rast(water_u, tsel_c)))
#     wat_arrow$u <- as.vector(terra::values(terra::rast(water_u, tsel_c)))
#     wat_arrow$v <- as.vector(terra::values(terra::rast(water_v, tsel_c)))
# 
#     scale <- 0.05
#     aspect_ratio <- 1 / cos(mean(plot_ylim) * pi / 180)
#     arrows(
#       wat_arrow$x, wat_arrow$y,
#       wat_arrow$x + scale * aspect_ratio * wat_arrow$u,
#       wat_arrow$y + scale * wat_arrow$v,
#       length = 0.01, col = "lightpink"
#     )
# 
#     win_arrow <- as.data.frame(terra::crds(terra::rast(wind_u, tsel_w)))
#     win_arrow$u <- as.vector(terra::values(terra::rast(wind_u, tsel_w)))
#     win_arrow$v <- as.vector(terra::values(terra::rast(wind_v, tsel_w)))
# 
#     arrows(
#       win_arrow$x, win_arrow$y,
#       win_arrow$x + scale * aspect_ratio * win_arrow$u,
#       win_arrow$y + scale * win_arrow$v,
#       length = 0.01, col = "lightblue"
#     )
# 
#     lapply(gd_load, plot, add = TRUE, col = "lightgrey")
#     text(2.5, 52.9, "r-coders-anonymous.blogspot.com", adj = c(0, 0.5), cex = 1.5)
#     axis(1)
#     axis(2)
#     dev.off()
#   }
# }
```












# ############################################################## #

# Old version ---

```{r func}
## function that reads netcdf files and puts them in a workable stack object:
read.mover <- function(ncdfile, xvar = "u", yvar = "v") {
  u <- stack(ncdfile, varname = xvar)
  v <- stack(ncdfile, varname = yvar)
  ## Get time indices:
  z_time      <- attr(u, "z")

  if (class(z_time[[1]]) == "character") {
    
    z_time <- as.POSIXct(z_time[[1]], tz = "UTC")
    
  } else {
    
    z_time_info <- tolower(strsplit(tolower(names(z_time)), " since ")[[1]])
    z_time      <- z_time[[1]] * switch(
      z_time_info[1], 
      seconds = 1, 
      minutes = 60, 
      hours   = 60 * 60, 
      days    = 60 * 60 * 24
      )
    z_time      <- as.POSIXct(
      z_time, 
      origin = as.POSIXct(z_time_info[2], tz = "UTC"), 
      tz = "UTC"
      )
    rm(z_time_info)
  }

  return(list(u = u, v = v, z = z_time))
}

## Countries for which the shorelines are downloaded and plotted
countries <- "NLD"
gd_load <- lapply(
  as.list(countries),
  function(x) {
    gadm <- ""
    lev <- 0
    if (x == "NLD") lev <- 1
    # try(gadm <- raster::getData("GADM", country = x, level = lev))

    gadm <- tryCatch(
      geodata::gadm(country = x, level = lev, path = tempdir()),
      error = function(e) NULL
    )

    if (class(gadm) == "SpatialPolygonsDataFrame" && lev == 1 && !is.character(gadm)) {
      gadm <- unionSpatialPolygons(gadm, gadm@data$ENGTYPE_1 != "Water body")[2]
    }
    gadm
  }
)

## function that extrapolates water currents for NA values
# extend.data <- function(r, band = 21) {
# extend.data <- function(r, cell_neighbor = 21) {
# 
#   ## replace NA values of raster data at coasts with nearest neighbors:
#   # fill.na <- function(x, i = round(band^2 / 2)) {
#   fill.na <- function(x, i = round(cell_neighbor^2 / 2)) {
#     if (is.na(x)[i]) {
#       return(mean(x, na.rm = TRUE))
#     } else {
#       return(x[i])
#     }
#   }
# 
#   r <- raster::setValues(
#     x = r,
#     values = raster::getValues(
#       raster::focal(
#         x     = r,
#         # w     = matrix(1, band, band),
#         w     = matrix(1, cell_neighbor, cell_neighbor),
#         fun   = fill.na,
#         pad   = TRUE,
#         na.rm = FALSE
#       )
#     )
#   )
#   r
# }

extend.data <- function(r, band = 21) {
  raster::focal(
    x        = r,
    w        = matrix(1, band, band),
    fun      = mean,
    na.rm    = TRUE,
    pad      = TRUE,
    padValue = NA,
    NAonly   = TRUE
  )
}

# plot(raster(water_u, tsel_c))
# plot(extend.data(raster(water_u, tsel_c)))
# plot(extend.data2(raster(water_u, tsel_c)))
```

## STEP 1: load the water currents and wind data

```{r}
# Read the water current data
temp <-
  here(path_save) %>%
  fs::dir_ls(regexp = dataset_id_phy) %>%
  .[[1]] %>%
  read.mover("uo", "vo")
water_u    <- temp$u
water_v    <- temp$v
water_time <- temp$z
rm(temp)

## Read the wind data:
temp <- 
  here(path_save) %>%
  fs::dir_ls(regexp = dataset_id_wind) %>%
  .[[1]] %>%
  read.mover("northward_wind", "eastward_wind")
wind_u     <- temp$u
wind_v     <- temp$v
wind_time  <- temp$z
rm(temp)



# ## Read the water current data:
# temp       <- read.mover("input/MetO-NWS-PHYS-hi-Agg-all-levels_1468581872238.nc",
#                          "vozocrtx", "vomecrty")
# water_u    <- temp$u
# water_v    <- temp$v
# water_time <- temp$z
# rm(temp)
# 
# ## Read the wind data:
# temp       <- read.mover("input/CERSAT-GLO-BLENDED_WIND_L4-V5-OBS_FULL_TIME_SERIE_1468508078544.nc",
#                          "eastward_wind_rms", "northward_wind_rms")
# wind_u     <- temp$u
# wind_v     <- temp$v
# wind_time  <- temp$z
# rm(temp)
```


## STEP 2: specify particle release details


```{r}
## time at which the simulated release should start:
release.start <- as.POSIXct("2018-01-02 0:00", tz = "UTC")

## time at which the simulated release should end:
release.end   <- as.POSIXct("2018-01-03 00:00", tz = "UTC")

## number of particles to release in total:
# release.n     <- 5000
release.n     <- 10

## release lon-coordinate:
release.x     <- 3.7

## release lat-coordinate:
release.y     <- 52

## release depth (not used yet in simulation)
release.z     <- 0

```

## STEP 3: specify simulation details

```{r}
## time at which the simulation should start:
simulation.start.time <- as.POSIXct("2018-01-02 06:00", tz = "UTC")

## time step of simulation in seconds:
# simulation.time.step <- 60 * 60 / 4 # every 15 min
simulation.time.step <- 60 * 60 * 6 # every 6 hours

## integer value indicating every n-th time step that should be exported:
simulation.out.step   <- 4

## time at which the simulation should stop
simulation.end.time   <- as.POSIXct("2018-01-10 06:00", tz = "UTC")

## create a time series based on the start and end time and the time step:
simulation.time       <- seq(
  simulation.start.time, 
  simulation.end.time - simulation.time.step, 
  simulation.time.step
  )

## for each time step determine how many particles are released:
n.particles <- rep(NA, length(simulation.time))

# if simulation is before start release
n.particles[simulation.time <= release.start] <- 0

# if simulation is after end of release
n.particles[simulation.time >= release.end]   <- release.n

n.particles[is.na(n.particles)] <- approx(
  c(0, release.n), 
  n = 2 + sum(is.na(n.particles))
  )$y[c(-1, -(2 + sum(is.na(n.particles))))]

n.particles <- round(n.particles)

## create a SpatialPointsDataFrame for each release of the particles:
## number of released particles 
particles <- data.frame(
  pos.x    = rep(release.x, release.n), # lon
  pos.y    = rep(release.y, release.n), # lat
  pos.z    = rep(release.z, release.n)  # depth
  )
coordinates(particles) <- ~ pos.x + pos.y
proj4string(particles) <- CRS("+proj=longlat +datum=WGS84")

## create a list in which results should be stored:
part.results          <- list()
previous.position     <- particles

## projection to be used to calculate particle displacement in meters:
part.proj <- CRS("+proj=utm +zone=31 +datum=WGS84")

## drift factor indicates with which fraction particles are displaced by the wind speed:
drift.factor    <- 0.03

## the diffusion coefficient, used in random walk simulation
diffusion.coeff <- 1e5
```

## STEP 4: loop all simulation time steps

```{r}
out_list <- 1

## start looping the simulation times
for (i in 1:length(simulation.time)) {
  print(i)
  
  ## Get the water currents that are closest to the current simulation time
  # tdif   <- abs(as.numeric(water_time) - as.numeric(simulation.time[i]))
  tdif   <- abs(water_time - simulation.time[i])
  tsel_c <- which(tdif == min(tdif))[[1]] # <-- closest time to the simulation time step
  twat_u <- extend.data(raster(water_u, tsel_c))
  twat_v <- extend.data(raster(water_v, tsel_c))
  
  # tdif   <- abs(as.numeric(wind_time) - as.numeric(simulation.time[i]))
  tdif   <- abs(wind_time - simulation.time[i])
  tsel_w <- which(tdif == min(tdif))[[1]] # <-- closest time to the simulation time step
  twin_u <- extend.data(raster(wind_u, tsel_w))
  twin_v <- extend.data(raster(wind_v, tsel_w))

  ## Get the last known particle locations and transform coordinates into meters:
  # converts decimal degree into meters as matrix x, y coords
  part.coords <- coordinates(spTransform(previous.position, part.proj))

  ## displace particles with with water_currents
  part.coords[0:n.particles[i],] <- 
    part.coords[0:n.particles[i],] + # extract number of coords
    # add u and v to particle and number of seconds
    # m/s * s = m in u and v
    simulation.time.step *
    cbind(
      raster::extract(twat_u, previous.position, "bilinear"),
      raster::extract(twat_v, previous.position, "bilinear")
      )[0:n.particles[i],]
  
  ## displace particles  with wind
  # same as above but with wind and drift factor
  part.coords[0:n.particles[i],] <- 
    part.coords[0:n.particles[i],] +
    # add u and v from wind with a less movement from wind (drift factor)
    simulation.time.step * drift.factor *
    cbind(
      raster::extract(twin_u, previous.position, "bilinear"),
      raster::extract(twin_v, previous.position, "bilinear")
      )[0:n.particles[i],]
  
  ## displace with random walk (diffusion):
  part.coords[0:n.particles[i],] <- 
    part.coords[0:n.particles[i],] +
    cbind(
      runif(n.particles[i], -1, 1),
      runif(n.particles[i], -1, 1)
      ) * 
    sqrt(diffusion.coeff * simulation.time.step * 6 / 1e4)

  ## turn particle information into a SpatialPointsDataFrame object:
  part.coords <- as.data.frame(part.coords)
  # coordinates(part.coords) <- ~ pos.x + pos.y # <---
  coordinates(part.coords) <- ~ coords.x1 + coords.x2

  # add back CRS because when convert to dataframe, will loose it
  proj4string(part.coords) <- part.proj
  
  ## back-transform coordinates to original projection:
  part.coords <- spTransform(part.coords, CRS(proj4string(previous.position)))

  ## store the new locations in the results list:
  previous.position     <- part.coords

  
  
  ## for each n-th simulation step store the results (where n = simulation.out.step):
  # if ((i - 1) %% 4 == 0) {
  if ((i - 1) %% simulation.out.step == 0) {
    
    ## add an indicator of the release state of the particles:
    part.coords <- SpatialPointsDataFrame(
      part.coords, 
      data.frame(released = rep(FALSE, release.n))
      )
    
    part.coords@data$released[0:n.particles[i]] <- TRUE
    # part.results[[(i - 1)/simulation.out.step + 1]] <- part.coords
    part.results[[out_list]] <- part.coords

    ## name the list element after the time index it represents:
    # names(part.results)[(i - 1)/simulation.out.step + 1] <- paste(
    names(part.results)[out_list] <- paste(
      format(simulation.time[i], "%Y-%m-%d %H:%M:%S"),
      attr(as.POSIXlt(simulation.time[i]),"tzone")
      )

    plot.lims <- apply(coordinates(part.coords), 2, mean)
    plot.ylim <- c("ymin" = plot.lims[[2]] - 2, "ymax" = plot.lims[[2]] + 2)
    plot.xlim <- c("xmin" = plot.lims[[1]] - 2, "xmax" = plot.lims[[1]] + 2)
    # plot.ylim <- c(part2d_bbox["ymin"], part2d_bbox["ymax"])
    # plot.xlim <- c(part2d_bbox["xmin"], part2d_bbox["xmax"])
    
    scale <- 1
    aspect.ratio <- 1/cos(mean(plot.ylim)*pi/180)
    
    ## extract water current
    wat_arrow   <- as.data.frame(coordinates(raster(water_u, tsel_c))) # coord
    wat_arrow$u <- as.vector(t(as.matrix(raster(water_u, tsel_c))))    # u 
    wat_arrow$v <- as.vector(t(as.matrix(raster(water_v, tsel_c))))    # v
    
    # filter to percent of rows
    wat_arrow <-  wat_arrow[seq(1, nrow(wat_arrow), by = 4), ]
    wat_arrow <- wat_arrow[!(is.na(wat_arrow$u) & is.na(wat_arrow$v)),]
    
    ## extract wind current
    win_arrow   <- as.data.frame(coordinates(raster(wind_u, tsel_w))) # coord
    win_arrow$u <- as.vector(t(as.matrix(raster(wind_u, tsel_w))))    # u 
    win_arrow$v <- as.vector(t(as.matrix(raster(wind_v, tsel_w))))    # v
    
    win_arrow <- win_arrow[seq(1, nrow(win_arrow), by = 4), ]
    win_arrow <- win_arrow[!(is.na(win_arrow$u) & is.na(win_arrow$v)),]
    
    # scale <- 0.05
    # aspect.ratio <- 1/cos(mean(plot.ylim)*pi/180)
    
    par(mar = c(3, 3, 2, 1))
    
    # plot points
    png(
      # here(plot_out_path, sprintf("out%03d.png", (i - 1)/4)),
      here(plot_out_path, sprintf("out%03d.png", out_list)),
      width = 2000, 
      height = 1500,
      # width  = 640,
      # height = 480,
      res    = 300,
      type   = "cairo"
      )
    
    plot(
      # part.results[[(i - 1)/4 + 1]], 
      part.results[[out_list]], 
      pch  = "+", 
      # xlim = c(3, 5),
      xlim = plot.xlim, 
      ylim = plot.ylim,
      col = "red",
      cex = 3,
      main = names(part.results)[out_list]
      # main = paste(
        # format(simulation.time[i], "%Y-%m-%d %H:%M:%S"),
        # attr(as.POSIXlt(simulation.time[i]),"tzone"))
      )
    
    # water arrow
    arrows(
      wat_arrow$x, 
      wat_arrow$y,
      wat_arrow$x + scale * aspect.ratio * wat_arrow$u,
      wat_arrow$y + scale * wat_arrow$v, 
      length = 0.01, 
      col = "lightpink"
      )
    
    # wind arrow
    arrows(
      win_arrow$x, 
      win_arrow$y,
      win_arrow$x + 0.025 * aspect.ratio * win_arrow$u,
      win_arrow$y + 0.025 * win_arrow$v, 
      length = 0.01, 
      col = "lightblue"
      )
    
    ## draw coastlines:
    if (is.null(gd_load)) {
      lapply(gd_load, plot, add = TRUE, col = "lightgrey")
    }
    
    # text(2.5, 52.9, "r-coders-anonymous.blogspot.com", adj = c(0, 0.5), cex = 1.5)
    axis(1)
    axis(2)
    dev.off()
    
    out_list <- out_list + 1
  }
  
}
```


```{r}
## save the resulting particle information:
save(part.results, file = here(plot_out_path, "..", "out.RData"), compress = TRUE)
```

